Multi-Core Model Checking and Maximum Satisfiability Applied to Hardware-Software Partitioning

### 2015-09-28 Docs/TODAES/paper.tex
review section 3

The informal model can be described by five characteristics. First, there is only one software context, i.e., there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning ~\cite{Teich2012}.

@@

\vspace{3 mm}
Three different optimization and decision problems can be defined. In this paper, the focus is on the case that $ S_0 $ is given, i.e., to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal ({\it i.e.}, system with hard real-time constraints). Based on Equations~\ref{linear-programming-problem} and~\ref{software-communication-costs}, the constraints can be reformulated as

@@

%\comment{
%----------------------------------------------
%\section{Partitioning problem using ILP-based,Genetic Algorithms}
%\label{ILPGA}
%----------------------------------------------

%The ILP and GA were taken from our previous studies~\cite{Trindade2015,Trindade2016}. Both use slack variables in order to be %possible to represent constraints and to use commercial tools. However, GA had improvements from the parameters of related %studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and %resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB~\cite{TheMathWorks2013}: the %population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of %generations changed from $100*NumberOfVariables$ (default) to $75$.
%}

%----------------------------------------------
\section{Analysis of the partitioning problem using ESBMC}
\label{Analysis-of-the-partitioning-problem-using-ESBMC}
%----------------------------------------------

As computer hardware architecture moves from single- to multi-cores, parallel programming environments should be exploited to take advantage of the ability to run several threads on different processing cores. This section describes the verification algorithm using sequential ESBMC, followed by three multi-core model checking algorithms and the integration of the MaxSMT solver $\nu$Z, in order to speed up the HW-SW partitioning verification.

@@

It is possible to inform to ESBMC with which type of values the variables must be tested. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. Those values that change for each test will generate a possible solution and obey the constraints. If this is achieved, then a feasible solution is found and the ASSUME directive is responsible for ensuring the compliance of constrains (i.e., $A.x \leq b$).

@@

To optimize the CPU resources utilization without modifying the underlying SMT Solver, the Open Multi-Processing (OpenMP) library~\cite{Dagum1998} is used in this present work as a front-end for ESBMC. Fig.~\ref{ESBMC-Multi-core} shows our first approach called ``Multi-core ESBMC''.

@@

ESBMC-SS obtains the problem specification represented by a C program. The HW-SW partitioning is violated, when the correct optimum value (\textit{TipH}) parameter is reached; ESBMC-SS starts a parallel region with different instances of ESBMC, based on the number of available processing cores. All these ESBMC instances run independently of each other, as shown in Fig.~\ref{ESBMC-Multi-core}. Note that there is no shared-memory (or message-passing) mechanism among the threads. In particular, different threads are managed by the OpenMP API, which is responsible for the thread lifecycle: start, running, and dead states, using different values as condition. After executing $N$ instances, if there is no code violation, then multi-core ESBMC starts new instances again. During the parallel region execution, if a violation is found in any running thread, then it presents a counterexample with the violation condition and the verification time. If all threads of the batch processing are terminated, then ESBMC-SS finishes its execution.

@@

The previous parallelization is implemented by continuously forking ESBMC instances until the first violation is found. However, since OpenMP only returns from a parallelized loop, when every forked thread finishes, some processing cores could remain idle for some period of time.

@@

The most optimized approach applies a parallelized binary search to reduce the amount of steps to be executed in order to find the optimal solution. A controller is designed to return the step to be executed so that the number of verification runs are substantially reduced. The parallelized binary search accomplishes this by splitting the domain of possible values into intervals and then by returning the middle of the largest interval so that two new intervals are created.

@@

As an example, given a problem of domain from $1$ to $20$ (see Fig.~\ref{Binary-Step-Calculation}), we firstly create an initial interval from $1$ to $20$. When the next available core asks for a step to be executed, the controller obtains the largest interval, {\it i.e.}, $\left[1,20\right]$, divides it by two, which creates two new intervals ({\it i.e.}, $\left[1,9\right]$ and $\left[11,20\right]$), and returns the middle of the original interval ({\it i.e.}, $10$). The controller also checks whether an interval has less than two elements to avoid creating empty or invalid intervals.

@@

Note that there might gaps between steps, which are produced by the customized binary search. For instance, in the example shown in Fig.~\ref{Binary-Step-Calculation}, if step $10$ returns \textit{false}, then one can conclude that all steps after $10$ is \textit{false} as well. However, if the same step $10$ returns \textit{true}, we can assume that all steps before $10$ is \textit{true} as well. As a result, an auxiliary method to remove unnecessary steps is implemented in the controller by removing or shrinking existing intervals. This approach leads to a high impact in the verification time. However, if a step is running and is not needed anymore, the worker kills the forked process and starts a new one.

@@

\vspace{3 mm}
Note that Algorithm~\ref{Steps-Calculation-using-Intervals} is called from each worker in order to get the next step to execute if it exists; otherwise, either zero or a negative number is returned.
From lines $4$ to $9$, the algorithm finds the largest interval. Then, from the line $10$ the largest interval is removed and the median is calculated in line $11$. After that, two new intervals are created, the left side (in line $14$) and the right side (in line $16$). At the end, the median is returned.

>>>>

The informal model can be described by five characteristics. First, there is only one software context, {\it i.e.}, there is just one general-purpose processor, and there is only one hardware context. The components of the system must be mapped to either one of these two contexts. Second, the software implementation of a component is associated with a software cost, which is the running time of the component. Third, the hardware implementation of a component has a hardware cost, which can be area, heat dissipation, and energy consumption. Fourth, based on the premise that hardware is significantly faster than software, the running time of the components in hardware is considered as zero. Finally, if two components are mapped to the same context, then there is no overhead of communication between them; otherwise, there is an overhead. The consequence of these assumptions is that scheduling does not need to be addressed in this work. Hardware components do not need scheduling, because the running time is assumed to be zero. Because there is only one processor, software components do not need to be scheduled as well. Therefore, the focus is only on the partitioning problem. That configuration describes a first-generation co-design, where the focus is on bipartitioning ~\cite{Teich2012}.

@@

\vspace{3 mm}
Three different optimization and decision problems can be defined. In this paper, the focus is on the case that $ S_0 $ is given, {\it i.e.}, to find a $P$ HW-SW partitioning so that $ S_p \leq S_0 $ and $ H_p $ is minimal, which is thus related to system with hard real-time constraints. Based on Equations~\ref{linear-programming-problem} and~\ref{software-communication-costs}, the constraints can be reformulated as

@@

%----------------------------------------------
\section{Analysis of the partitioning problem}
\label{Analysis-of-the-partitioning-problem-using-ESBMC}
%----------------------------------------------

As computer hardware architecture moves from single- to multi-cores, parallel programming environments should be exploited to take advantage of the ability to run several threads on different processing cores. This section describes the verification algorithm using sequential ESBMC, followed by three multi-core model checking algorithms and the integration of the MaxSMT solver $\nu$Z into ESBMC, in order to speed up the HW-SW partitioning verification. HW-SW partitioning using ILP-based and Genetic Algorithms are also explained.

%----------------------------------------------
\subsection{Partitioning problem using ILP-based,Genetic Algorithms}
\label{ILPGA}
%----------------------------------------------

The ILP and GA were taken from our previous studies~\cite{Trindade2015,Trindade2016}. Both use slack variables in order to be possible to represent constraints and to use commercial tools. However, GA had improvements from the parameters of related studies in order to increase the solution accuracy without producing timeout. The tuning was performed by empirical tests and resulted in changing of three parameters, which are passed to the function \textit{ga} of MATLAB~\cite{TheMathWorks2013}: the population size was set from $300$ to $500$, the Elite count changed from $2$ (default value) to $50$, and the number of generations changed from $100*NumberOfVariables$ (default) to $75$.

@@

It is possible to inform to ESBMC with which type of values the variables must be tested. Therefore, there is a declaration to populate all decision variables $x$ with non-deterministic Boolean values. Those values that change for each test will generate a possible solution and obey the constraints. If this is achieved, then a feasible solution is found and the ASSUME directive is responsible for ensuring the compliance of those constrains (i.e., $A.x \leq b$).

@@

To optimize the CPU resources utilization without modifying the underlying SMT solver, the Open Multi-Processing (OpenMP) library~\cite{Dagum1998} is used in this present work as a front-end for ESBMC. Fig.~\ref{ESBMC-Multi-core} shows our first approach called ESBMC sequential-search ``ESBMC-SS''.

@@

ESBMC-SS obtains the problem specification represented by a C program. The HW-SW partitioning is violated, when the correct optimum value (\textit{TipH}) parameter is reached; ESBMC-SS starts a parallel region with different instances of ESBMC, based on the number of available processing cores. All these ESBMC instances run independently of each other, as shown in Fig.~\ref{ESBMC-Multi-core}. Note that there is no shared-memory (or message-passing) mechanism among threads. In particular, different threads are managed by the OpenMP API, which is responsible for the thread life-cycle: start, running, and dead states, using different values as condition. After executing $N$ instances, if there is no code violation, then ESBMC-SS starts new instances again; this represents a sequential-search on a multi-core environment. During the parallel region execution, if a violation is found in any running thread, then it presents a counterexample with the violation condition and the verification time. If all threads of the batch processing are terminated, then ESBMC-SS finishes its execution.

@@

The previous parallelization is implemented by continuously forking ESBMC instances in a sequential manner until the first violation is found. However, since OpenMP only returns from a parallelized loop, when every forked thread finishes, some processing cores could remain idle for some period of time.

@@

The most optimized approach applies a parallelized binary-search to reduce the amount of steps to be executed in order to find the optimal solution. A controller is designed to return the step to be executed so that the number of verification runs are substantially reduced. The parallelized binary search accomplishes this by splitting the domain of possible values into intervals and then by returning the middle of the largest interval so that two new intervals are created.

@@

As an example, given a problem of domain from $1$ to $20$ (see Fig.~\ref{Binary-Step-Calculation}), we firstly create an initial interval from $1$ to $20$. When the next available core requests a step to be executed, the controller obtains the largest interval, {\it i.e.}, $\left[1,20\right]$, divides it by two, which creates two new intervals ({\it i.e.}, $\left[1,9\right]$ and $\left[11,20\right]$), and returns the middle of the original interval ({\it i.e.}, $10$). The controller also checks whether an interval has less than two elements to avoid creating empty or invalid intervals.

@@

Note that there might gaps between steps, which are produced by the customized binary-search. For instance, in the example shown in Fig.~\ref{Binary-Step-Calculation}, if step $10$ returns \textit{false}, then one can conclude that all steps after $10$ is \textit{false} as well. However, if the same step $10$ returns \textit{true}, we can assume that all steps before $10$ is \textit{true} as well. As a result, an auxiliary method to remove unnecessary steps is implemented in the controller by removing or shrinking existing intervals. This approach leads to a high impact in the verification time. However, if a step is running and is not needed anymore, the worker kills the forked process and starts a new one.

@@

\vspace{3 mm}
Note that Algorithm~\ref{Steps-Calculation-using-Intervals} is called from each worker in order to get the next step to execute if it exists; otherwise, either zero or a negative number is returned.
From lines $4$ to $9$, the algorithm finds the largest interval. Then, from line $10$ the largest interval is removed and the median is calculated in line $11$. After that, two new intervals are created, the left side (in line $14$) and the right side (in line $16$). At the end, the median is returned.
